---
title: "Egypt Seshat Presentation"
author: "Corey White"
format: 
  beamer:
    theme: Berlin
    colortheme: dolphin
    aspectratio: 43
knitr:
  opts_chunk: 
    echo: true
    eval: true
    prompt: false
    tidy: false
    comment: NA
    message: false
    warning: false
    fig.align: 'center'
    cache: true
---

# \vspace*{\fill}

\begin{center}
Plotting Rise and Decline in Egypt's Old Kingdom
\end{center}
\vspace*{\fill}

# Predynastic to the the First Intermediate Period

## What does systemic decline look like in ancient times?

Ancient Egypt's Old Kingdom decline resulted in the rise of provincial governors, called *Nomarchs*, who dissected the kingdom. This data looks at Upper Egypt, the southern provinces farther from the capital in Memphis.

# 

### Over the course of the Old Kingdom, Upper Egypt's population steadily increased.

```{r echo = FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(haven)
Seshat <- read_excel('seshat.xlsx')
egypt <- Seshat |>
  filter(NGA == 'Upper Egypt')
egtest <- egypt |>
  filter(NGA == 'Upper Egypt', Polity == 'EgDyn1*' | Polity =='EgDyn2*' | Polity == 'EgOldK1' | Polity == 'EgOldK2' | Polity == 'EgRegns' | Polity == 'EgNaqa3') 
egtest2 <- egtest |>
  filter(Variable == 'Polity territory' | Variable == 'Polity Population' |Variable == 'Largest communication distance' | Variable == 'Fastest individual communication' | Variable == 'Population of the largest settlement' )

egdata <- egtest2 |>
  group_by(Polity, Variable)


mytheme <- theme(
  plot.background = element_rect(fill = '#F9F7F2', color = NA),
  panel.grid.major.y = element_line(colour = 'grey50', linetype = 'dotted'),
  panel.grid.major.x = element_blank(),
  axis.text = element_text(family = 'serif', color = '#4A4A4A'),
  plot.title = element_text(family = 'serif', face = 'bold', size = 16, color = '#2B4F60'),
  panel.border = element_rect(fill = NA, colour = 'black'),
  legend.position = 'none',
  panel.background = element_rect(fill = 'grey99')
)
```
```{r echo= FALSE}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 
popplot <- egdata |>
  mutate(Polity2 = factor(Polity, levels = c('EgNaqa3', 'EgDyn1*', 'EgDyn2*', 'EgOldK1', 'EgOldK2', 'EgRegns'))) |>
  filter(Variable == 'Polity Population') |>
  mutate(`Value From` = as.numeric(`Value From`)) |>
  ggplot(aes(x = Polity2, y = `Value From`, fill = Polity)) +
  geom_col(color = NA, size = .6) +
  labs(
    x = '(a) Total Polity Population',
    y = 'Maximum Population'
  ) +
  scale_fill_manual(
    values = c('#556B2F','#556B2F','#556B2F','#556B2F','#556B2F','#556B2F')
  )+
  scale_x_discrete(
    labels = c(
      'EgNaqa3' = 'Predynastic Period',
      'EgDyn1*' = 'Dynasty I',
      'EgDyn2*' = 'Dynasty II',
      'EgOldK1' = 'Early 
        Old Kingdom',
      'EgOldK2' = 'Late 
        Old Kingdom',
      'EgRegns' = 'First
      Intermediate
      Period'
    )
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, .01))) + mytheme

(popplot)

```

# 

## Centralization

The Old Kingdom centralized Egypt, and built Egypt's first true cities. Settlement population expanded.
Measurements form the First Intermediate period, however, show a decline in maximum settlement population

#

```{r echo = FALSE}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 
setpopplota <- egdata |>
  mutate(Polity2 = factor(Polity, levels = c('EgNaqa3', 'EgDyn1*', 'EgDyn2*', 'EgOldK1', 'EgOldK2', 'EgRegns'))) |>
  filter(Variable == 'Population of the largest settlement') |>
  mutate(`Value From` = as.numeric(`Value From`)) |>
  group_by(Polity2) |>
  summarise(frommax = max(`Value From`, na.rm = TRUE)) 

setpopplot <- 
  ggplot(setpopplota, aes(x = Polity2,  y = frommax, fill = Polity2)) +
  geom_col(color = NA, size = .6) +
  scale_y_continuous(
    labels = scales::label_comma(), 
    expand = expansion(mult = c(0, .01)))+
  labs(
    x = ' (b) Settlement Population',
    y = 'Population of the Largest Settlement'
  ) +
  scale_x_discrete(
    labels = c(
      'EgNaqa3' = 'Predynastic Period',
      'EgDyn1*' = 'Dynasty I',
      'EgDyn2*' = 'Dynasty II',
      'EgOldK1' = 'Early 
        Old Kingdom',
      'EgOldK2' = 'Late 
        Old Kingdom',
      'EgRegns' = 'First Intermediate
    Period'
    )
  ) +
  scale_fill_manual(
    values = c('#556B2F','#556B2F','#556B2F','#556B2F','#556B2F','#556B2F')
  )+ mytheme
(setpopplot)

```

#
 
## Territorial Extent
Lastly, we can see that the maximum extent of Egypt's polities decreased over time, indicating a breakdown of 
central authority.

# 
```{r echo = FALSE}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 

terrplota <- egdata |>
  mutate(Polity2 = factor(Polity, levels = c('EgNaqa3', 'EgDyn1*', 'EgDyn2*', 'EgOldK1', 'EgOldK2', 'EgRegns'))) |>
  filter(Variable == 'Polity territory') |>
  mutate(`Value From` = as.numeric(`Value From`)) |>
  group_by(Polity2) |>
  summarise(frommax = max(`Value From`, na.rm = TRUE)) 

terrplot <- 
  ggplot(terrplota, aes(x = Polity2,  y = frommax, fill = Polity2)) +
  geom_col(color = NA, size = .6) +
  scale_y_continuous(
    labels = scales::label_comma(), 
    expand = expansion(mult = c(0, .01)))+
  labs(
    x = '(c) Territorial Extent of Polities',
    y = 'Maximum Territory, in KM2'
  ) +
  scale_x_discrete(
    labels = c(
      'EgNaqa3' = 'Predynastic Period',
      'EgDyn1*' = 'Dynasty I',
      'EgDyn2*' = 'Dynasty II',
      'EgOldK1' = 'Early 
        Old Kingdom',
      'EgOldK2' = 'Late 
        Old Kingdom',
      'EgRegns' = 'First Intermediate
    Period'
    )
  ) +
  scale_fill_manual(
    values = c('#556B2F','#556B2F','#556B2F','#556B2F','#556B2F','#556B2F')
  )+ mytheme
terrplot

```

# 

## Synthesized Output

#

```{r echo = FALSE}
#| echo: false 
#| fig.height: 6
#| fig.width: 12

egcombine <- popplot + setpopplot + terrplot +
  plot_annotation(
    title = 'Population and Territory in Ancient Egypt', 
    subtitle = 'From Predynastic Egypt to the First Intermediate Period (3,300-2,055 BCE)',
    theme = theme(
      plot.title = element_text(size = 18, face = 'bold', hjust =.0),
      plot.subtitle = element_text(face = 'italic', hjust = .1)
    )
    
  )
(egcombine)
```

#

## Analysis
We can see that increased rurification combined with increased total population
signals a breakdown in central authority and command in Upper Egypt. Please consult the `README`
for a complete breakdown and analysis.