---
title: "Malagasy Opinions on Public Services and Servants"
subtitle: 'Corruption and Decline vs Violence Prevention'
author: "Corey White"
format: 
  beamer:
    theme: Singapore
    colortheme: seahorse
    aspectratio: 43
knitr:
  opts_chunk: 
    echo: true
    eval: true
    prompt: false
    tidy: false
    comment: NA
    message: false
    warning: false
    fig.align: 'center'
    cache: false
---

#  \vspace*{\fill}
\begin{center}
How do Malagasy Feel about Public Officials' Provision of Services?
\end{center}
\vspace*{\fill}

```{r echo = FALSE}
library(knitr)
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
mytheme <- theme(
      plot.background = element_rect(fill = "white", color = NA),
      panel.grid.major.x = element_blank(),
      axis.title = element_text(face = "bold", size = 10),
      axis.text.x = element_text(angle = 35, hjust = 1),
      plot.title = element_text(face = "bold", size = 12, color = "#213159"),
      panel.border = element_blank(),
      axis.line.x = element_line(color = "black")
    )

```

```{r echo=FALSE}
mad1 <- read_excel('mad1.xlsx', col_types = "text")

df <-mad1
df4 <- mad1 |>
  select(Q48A, starts_with('Q47')) |>
  mutate(across(everything(), ~ if_else(str_detect(., "Ne sait pas"), NA_character_, as.character(.)))) |>
  filter(!is.na(Q48A)) |>
  mutate(Q48A = fct_relevel(Q48A, 
                            "Tout à fait d’accord", 
                            "D’accord", 
                            "En désaccord", 
                            "Tout à fait en désaccord",
                            "Refuse de répondre"))
df5 <- df4 |>
  filter(!is.na(Q47K))|>
  filter(!is.na(Q47J)) |>
  filter(!is.na(Q47M)) |>
  filter(Q47J != "Refuse de répondre") |>
  mutate(Q47K = fct_relevel(Q47K,
                            'Très mal',
                            'Plutôt mal',
                            'Plutôt bien',
                            'Très bien'
                            )) |>
  mutate(Q47J = fct_relevel(Q47J,
                            'Très mal',
                            'Plutôt mal',
                            'Plutôt bien',
                            'Très bien',
                            'Refuse de répondre')) |>
  mutate(Q47M = fct_relevel(Q47M,
                            'Très mal',
                            'Plutôt mal',
                            'Plutôt bien',
                            'Très bien'))
```

# Core Findings

*Malagasy have an unfavorable opinion about the provision and maintenance of services on the island as well as public corruption, yet maintain a favorable view of the President and Assembly.*

# Findings

This dataset uses the Afrobarometer data on Madagascar. Key measurements include:

\begin{center}
Entretien des routes et ponts/Maintenance of roads and bridges



Lutte contre la corruption au sein de l’administration publique/Combating corruption in public administration


Prévention ou résolution des conflits violents/Prevention or resolution to violent conflicts


Performance du Président/Performance of the President
\end{center}

# Data


## Malagasy have an unfavorable view on the maintenance of roads and bridges
```{r}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 
#| fig-cap: 'Maintenance of roads and bridges'

df5|>
  ggplot(aes(x = Q47K, fill = Q47K)) +
  geom_bar(na.rm = T)+
labs(
  x = 'Opinion',
  y =' Number of Responses'
) +
  scale_fill_manual(
    values = c("Très mal" = "firebrick",
               "Plutôt mal" = "#E2725B",
               "Plutôt bien" = "darkseagreen",
               "Très bien" = "seagreen")
  )+ mytheme +
  scale_y_continuous(
    expand = expansion(mult = c(0, .01)))
```



# Malagasy have an unfavorable view on controlling public corruption
```{r}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 
#| fig-cap: 'Combating corruption in public administration'

df5|>
  ggplot(aes(x = Q47J, fill = Q47J)) +
  geom_bar(na.rm = T)+
  labs(
    x = 'Opinion',
    y =' Number of Responses'
  ) +
  scale_fill_manual(
    values = c("Très mal" = "firebrick",
               "Plutôt mal" = "#E2725B",
               "Plutôt bien" = "darkseagreen",
               "Très bien" = "seagreen")
  )+ mytheme +
  scale_y_continuous(
    expand = expansion(mult = c(0, .01)))
```

# Yet, Malagasy have a Favorable View of the President.
```{r}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 

df4 |> ggplot(aes(x =Q48A)) + geom_bar(na.rm = T, fill = 'salmon')+
  labs( title = 'Performance of the President', x = 'Performance', y = 'number of responses' ) + mytheme + coord_flip()
```


# Why do Malagasy favor the President if they cannot provide services?


We can determine this by looking at the violent conflict data
```{r}
#| echo: false 
#| fig.height: 3
#| fig.width: 5.5 
#| fig-cap: 'Prevention or resolution to violent conflicts'
df5|>
  ggplot(aes(x = Q47M, fill = Q47M)) +
  geom_bar(na.rm = T)+
  labs(
    x = 'Opinion',
    y =' Number of Responses'
  ) +
  scale_fill_manual(
    values = c("Très mal" = "firebrick",
               "Plutôt mal" = "#E2725B",
               "Plutôt bien" = "darkseagreen",
               "Très bien" = "seagreen")
  )+ mytheme +
  scale_y_continuous(
    expand = expansion(mult = c(0, .01)))
```
# Conclusion

We may be able to conclude about why Malagasy favor their administration despite corruption and the ill
maintenance of public services. If the Malagasy government is successful at ending violent conflicts,
citizens may be more likely to overlook their outright corruption and the decline of the economy
and public services including roads and electricity. To many Malagasy, outright stability is more
important than the provision of services by the government.

